Numerical Determination of the Hartree-Fock Stability of the Paramagnetic Electron Gas in 1, 2 and 3 Dimensions

### 2018-04-02 doc/papers/paper_skeleton/paper_skeleton.tex
Added comments for original draft

  Where occupied states are labelled by $i$ and $j$, virtual by $a$ and $b$, and $i \rightarrow a$ refers to the excitation which replaces state $i$ with state $a$. The relevant two electron notation is
  \begin{equation}
    \left< pq||rs \right> = \left< pq|rs \right> - \left< pq|sr \right>,
  \end{equation}
  where, 
  \begin{equation}
    \left< pq|rs \right> = \int \int \phi^*_p(\vec{r}_1) \phi^*_q(\vec{r}_2) \frac{1}{r_{12}} \phi_r(\vec{r}_1) \phi_s(\vec{r}_2) d\vec{r_1} d\vec{r_2}, 
  \end{equation}
  and states labeled with $p$, $q$, $r$, or $s$ are either occupied or virtual. For a paramagnetic HF solution (RHF), the Orbital Hessian factorizes into the matrices known to chemists as the singlet and triplet instability matrices ($\bf{{}^1H}$ and $\bf{{}^3H}$, respectively)\cite{Dunning1967,Seeger1977}.
  \begin{subequations}
    \begin{eqnarray}
      {}^{1}{A'}_{i\rightarrow a, j\rightarrow b} &=& 
      (\epsilon_a-\epsilon_i)\delta_{ij}\delta_{ab} 
      + 2\braket{aj|ib}-\braket{aj|bi}\\
      {}^{3}{A'}_{i\rightarrow a, j\rightarrow b} &=& 
      (\epsilon_a-\epsilon_i)\delta_{ij}\delta_{ab} 
      - \braket{aj|bi}\\
      {}^{1}{B'}_{i\rightarrow a, j\rightarrow b} &=& 2\braket{ab|ij}-\braket{ab|ji}\\
      {}^{3}{B'}_{i\rightarrow a, j\rightarrow b} &=& -\braket{ab|ji}
    \end{eqnarray}
  \end{subequations}
  The lowest eigenvalue of both of these matrices will reveal the stability of the RHF solution. If  ${}^1\bf{H}$ has negative eigenvalues, the solution is unstable with respect to changes that preserve singlet character (singlet instability), while if ${}^3\bf{H}$ has negative eigenvalues, the solution is unstable towards displacements that would change the wavefunction towards a triplet state (triplet instability). The solution is stable if neither ${}^1\bf{H}$ nor ${}^3\bf{H}$ has a negative eigenvalue. An eigenvalue of 0 does not indicate instability, and this case is discussed in detail by Cui et al \cite{Cui2013}.

@@

  A basis of plane waves with simple cubic symmetry was used for all calculations, and the number of k-points was incremented until the lowest eigenvalues were sufficiently converged {\color{red} (for 1D and 2D more data can be gathered but in 3D getting more data is computationally unfeasible at the moment)}. In the 1D calculations, a potential of $V_0 = 1$ in atomic units was used.
  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../images/singlet-stability-convergence.eps}
    \caption{The lowest eigenvalue of ${}^1\mathbf{H}$ as a function of $r_s$ with increasing number of k-points in 1, 2 and 3 dimensions.}
    \label{fig:singlet_convergence}
  \end{figure}
  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../images/triplet-stability-convergence.eps}
    \caption{The lowest eigenvalue of ${}^3\mathbf{H}$ as a function of $r_s$ with increasing number of k-points in 1, 2 and 3 dimensions.}
    \label{fig:triplet_convergence}
  \end{figure}
  To determine the $r_s$ at which the lowest eigenvalue of $\mathbf{H}$ changes sign, the lowest eigenvalues were linearly interpolated  {\color{red} (this can easily be changed to one of [(‘linear’, ‘nearest’, ‘zero’, ‘slinear’, ‘quadratic’, ‘cubic’ where ‘zero’, ‘slinear’, ‘quadratic’ and ‘cubic’ refer to a spline interpolation of zeroth, first, second or third order)]} as a function of $r_s$ and solving
  \begin{equation} \label{eq:transition_rs}
    \lambda_{min}(r_s) = 0.
  \end{equation}
  The solution to eq. \ref{eq:transition_rs} is referred to as the transition $r_s$ and this quantity was determined while increasing the number of k-points in the simulation to convergence (see Fig. \ref{fig:onset}).
  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../images/combined_onset.eps}
    \caption{The value of $r_s$ at which the singlet (green) and triplet (blue) instability    occurs with increasing number of k-points in 1, 2 and 3 dimensions. In the 1D case (left), the singlet instability never appeared in the calculations performed. }
    \label{fig:onset}
  \end{figure}

>>>>

  Where occupied states are labelled by $i$ and $j$, virtual by $a$ and $b$, and $i \rightarrow a$ refers to the excitation which replaces state $i$ with state $a$. The relevant two electron notation is
  \begin{equation}
    \left< pq||rs \right> = \left< pq|rs \right> - \left< pq|sr \right>,
  \end{equation}
  where, 
  \begin{equation}
    \left< pq|rs \right> = 
    \int \int \phi^*_p(\vec{r}_1) \phi^*_q(\vec{r}_2) 
    V(\vec{r}_1, \vec{r}_2) 
    \phi_r(\vec{r}_1) \phi_s(\vec{r}_2) 
    d\vec{r_1} d\vec{r_2}, 
  \end{equation}
  and states labeled with $p$, $q$, $r$, or $s$ are either occupied or virtual. For a paramagnetic HF solution (RHF), the Orbital Hessian factorizes into the matrices known to chemists as the singlet and triplet instability matrices ($\bf{{}^1H}$ and $\bf{{}^3H}$, respectively)\cite{Dunning1967,Seeger1977}.
  \begin{subequations}
    \begin{eqnarray}
      {}^{1}{A'}_{i\rightarrow a, j\rightarrow b} &=& 
      (\epsilon_a-\epsilon_i)\delta_{ij}\delta_{ab} 
      + 2\braket{aj|ib}-\braket{aj|bi}\\
      {}^{3}{A'}_{i\rightarrow a, j\rightarrow b} &=& 
      (\epsilon_a-\epsilon_i)\delta_{ij}\delta_{ab} 
      - \braket{aj|bi}\\
      {}^{1}{B'}_{i\rightarrow a, j\rightarrow b} &=& 2\braket{ab|ij}-\braket{ab|ji}\\
      {}^{3}{B'}_{i\rightarrow a, j\rightarrow b} &=& -\braket{ab|ji}
    \end{eqnarray}
  \end{subequations}
  The lowest eigenvalue of both of these matrices will reveal the stability of the RHF solution. If  ${}^1\bf{H}$ has negative eigenvalues, the solution is unstable with respect to changes that preserve singlet character (singlet instability), while if ${}^3\bf{H}$ has negative eigenvalues, the solution is unstable towards displacements that would change the wavefunction towards a triplet state (triplet instability). The solution is stable if neither ${}^1\bf{H}$ nor ${}^3\bf{H}$ has a negative eigenvalue. An eigenvalue of 0 does not indicate instability, and this case is discussed in detail by Cui et al \cite{Cui2013}.

@@

  A basis of plane waves with simple cubic symmetry was used for all calculations, and the number of k-points was incremented until the lowest eigenvalues were sufficiently converged {\color{red} (for 1D and 2D more data can be gathered but in 3D getting more data is computationally unfeasible at the moment)}. In the 1D calculations, a potential of $V_0 = 1$ in atomic units was used.
  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../images/singlet-stability-convergence.pdf}
    \caption{The lowest eigenvalue of ${}^1\mathbf{H}$ as a function of $r_s$ with increasing number of k-points in 1, 2 and 3 dimensions.}
    \label{fig:singlet_convergence}
  \end{figure}
  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../images/triplet-stability-convergence.pdf}
    \caption{The lowest eigenvalue of ${}^3\mathbf{H}$ as a function of $r_s$ with increasing number of k-points in 1, 2 and 3 dimensions.}
    \label{fig:triplet_convergence}
  \end{figure}
  To determine the $r_s$ at which the lowest eigenvalue of $\mathbf{H}$ changes sign, the lowest eigenvalues were linearly interpolated  {\color{red} (this can easily be changed to one of [(‘linear’, ‘nearest’, ‘zero’, ‘slinear’, ‘quadratic’, ‘cubic’ where ‘zero’, ‘slinear’, ‘quadratic’ and ‘cubic’ refer to a spline interpolation of zeroth, first, second or third order)]} as a function of $r_s$ and solving
  \begin{equation} \label{eq:transition_rs}
    \lambda_{min}(r_s) = 0.
  \end{equation}
  The solution to eq. \ref{eq:transition_rs} is referred to as the transition $r_s$ and this quantity was determined while increasing the number of k-points in the simulation to convergence (see Fig. \ref{fig:onset}).
  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../images/combined_onset.pdf}
    \caption{The value of $r_s$ at which the singlet (green) and triplet (blue) instability    occurs with increasing number of k-points in 1, 2 and 3 dimensions. In the 1D case (left), the singlet instability never appeared in the calculations performed. }
    \label{fig:onset}
  \end{figure}
